---
title: "Epigenetic and MZT Biomarker Expression"
author: "Erin Chille"
date: "Last updated: 2020/11/09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Set up workspace

Remove prior objects and load necessary tools
```{r, warning=FALSE, message=FALSE}
rm(list=ls()) # removes all prior objects

library("RColorBrewer")
library("ggplot2")
library("tidyverse")
library("ggpubr")
library("DESeq2")
library("genefilter")
```


Load sample Information
```{r}
sample.info <- read.csv(file="Sample_Info/RNAseq_data.csv", header=T, sep=",") #load sample info
```

Load gene count Information
```{r}
counts <- as.data.frame(read.csv("1-QC-Align-Assemble/Output/gene_count_matrix.csv", row.names="gene_id"), colClasses = double) #Load normalized expression matrix from stringtie
```

#### Quality-filter gene counts  
Pre-filtering our dataset to reduce the memory size dataframe, increase the speed of the transformation and testing functions, and improve quality of statistical analysis by removing low-coverage counts. Removed counts could represent outliers in the data and removing these improves sensitivity of statistical tests. Here we will filter out the genes that are only present in fewer than two of the 24 ambient samples.
```{r}
#keep only ambient treatmentinfo and count data
dev <- c("AMB")
treatmentinfo_dev <- filter(sample.info, treatment %in% dev)
dim(treatmentinfo_dev) #rows should be 24

# delete sample columns corresponding to low and extreme low samples by mapping Null value to them
gcount_dev <- counts[treatmentinfo_dev$sample_id]
dim(gcount_dev) #columns should be 24

#create filter for the counts data
#gfiltdev <- rowSums(count(gcount_dev)) > 0
#set filter values for PoverA, P=100% percent of the samples have counts over A=10. This means that only 2 out of 24 (0.083) samples need to have counts over 10. Our smallest sample size for our life stages is two (fertilized egg, mid-gastrula, early-gastrula). By setting 2/24 as the P, this means if a particular gene is expressed only in 1 of these smallest life stages, it will be included in the analysis.
filt <- filterfun(pOverA(0.083,10))

#create filter for the counts data
gfiltdev <- genefilter(gcount_dev, filt)

#identify genes to keep by count filter
gkeepdev <- gcount_dev[gfiltdev,]


#identify genes to keep by count filter
gkeepdev <- gcount_dev[gfiltdev,]

#identify gene lists
gn.keepdev <- rownames(gkeepdev)

#gene count data filtered in PoverA, P percent of the samples have counts over A
gcount_filt_dev <- as.data.frame(gcount_dev[which(rownames(gcount_dev) %in% gn.keepdev),])

#How many rows do we have before and after filtering?
nrow(gcount_dev) #Before
nrow(gcount_filt_dev) #After
```

In order for the DESeq2 algorithms to work, the SampleIDs on the treatmentinfo file and count matrices have to match exactly and in the same order. The following R clump will check to make sure that these match.
```{r}
#Checking that all row and column names match. Should return "TRUE"
all(treatmentinfo_dev$sample_id %in% colnames(gcount_filt_dev))
all(treatmentinfo_dev$sample_id == colnames(gcount_filt_dev)) 
```

### Read normalization
We are now going normalize our read counts using VST-normalization in DESeq2

#### Construct the DESeq2 dataset

Merge the treatment and time_point columns into a new column , group. Set group as a factor.
```{r}
treatmentinfo_dev$time_point <- factor(treatmentinfo_dev$time_point, levels = c("Unfertilized_egg", "Fertilized_egg", "Cleavage", "Prawn_chip", "Early_gastrula", "Mid_gastrula", "Late_gastrula", "Planula", "Adult"))
```

Create a DESeqDataSet design from gene count matrix and labels. Here we set the design to look at time_point to test for any differences in gene expression across timepoints.
```{r}
#Set DESeq2 design
gdds_dev <- DESeqDataSetFromMatrix(countData = gcount_filt_dev,
                              colData = treatmentinfo_dev,
                              design = ~time_point)
```

#### Log-transform the count data
First we are going to log-transform the data using a variance stabilizing transforamtion (VST). This is only for visualization purposes. Essentially, this is roughly similar to putting the data on the log2 scale. It will deal with the sampling variability of low counts by calculating within-group variability (if blind=FALSE). Importantly, it does not use the design to remove variation in the data, and so can be used to examine if there may be any variability do to technical factors such as extraction batch effects.

To do this we first need to calculate the size factors of our samples. This is a rough estimate of how many reads each sample contains compared to the others. In order to use VST (the faster log2 transforming process) to log-transform our data, the size factors need to be less than 4. Otherwise, there could be artefacts in our results.
```{r}
SF.gdds_dev <- estimateSizeFactors(gdds_dev) #estimate size factors to determine if we can use vst  to transform our data. Size factors should be less than for to use vst
print(sizeFactors(SF.gdds_dev)) #View size factors
```

Our size factors are all less than 4, so we can use VST!
```{r}
gvst_dev <- vst(gdds_dev, blind=FALSE) #apply a variance stabilizing transforamtion to minimize effects of small counts and normalize wrt library size
head(assay(gvst_dev), 3) #view transformed gene count data
```
Find mean, median, mode of expression
```{r}
#Make a list of observations
exp_df <- as.data.frame(assay(gvst_dev))
exp_list <- c(exp_df$X119, exp_df$X120, exp_df$X121, exp_df$X127, exp_df$X132, exp_df$X153, exp_df$X154, exp_df$X159, exp_df$X162, exp_df$X163, exp_df$X167, exp_df$X179, exp_df$X180, exp_df$X184, exp_df$X212, exp_df$X215, exp_df$X218, exp_df$X221, exp_df$X359, exp_df$X361, exp_df$X375, exp_df$X1101, exp_df$X1548, exp_df$X1628)
sanity.check <- nrow(exp_df)*ncol(exp_df) #make sure num observations match
length(exp_list) == sanity.check #make sure num observations match. Should be true.

#Get stats and view distribution as a histogram
max(exp_list)
mean(exp_list)
median(exp_list)
min(exp_list)
hist(exp_list)
```


## Run BLAST search of biomarkers against Mcap protein fasta

Find orthologues of biomarkers within Mcap genome
```{bash}
#Make blast database from Mcap predicted protein fasta file
#/usr/local/ncbi/blast/bin/makeblastdb -in {path2ref}/Mcap.protein.fa -input_type fasta -dbtype prot -parse_seqids -out 2b-Epi-MZT-Enzyme-Expression/Input/Mcap.protein.blast.db

#Blast fasta file containing sequences of selected biomarkers against our sequences
#/usr/local/ncbi/blast/bin/blastp -query 2b-Epi-MZT-Enzyme-Expression/Input/cnidarian.Epi_MTZ_biomarkers.fa -db 2b-Epi-MZT-Enzyme-Expression/Input/Mcap.protein.blast.db -outfmt 6 -max_target_seqs 100 -evalue 1e-05 -out 2b-Epi-MZT-Enzyme-Expression/Output/Mcap_Epi_MZT_biomarker_hits_100.tsv

#Manually in text editor: (1.) Add column names: "queryprot", "gene_id", "pident", "length", "mismatch", "gapopen", "qstart", "qend", "sstart", "send", "evalue", "bitscore", and (2.) Remove ("augustus." and ".t1") from gene_ids using "command F" replace
#Manually in Excel: Add a column called Epi.gene that lists the alias for each of your target of enzymes. Check the annotation to see if there is any overlap in genes
```

Subset results for genes of interest
```{r}
hits <- read.table(file="2b-Epi-MZT-Enzyme-Expression/Output/cnidarian.Mcap_Epi_MZT_hits_100.csv", header=T, sep=",")
hits <- hits %>% distinct(Epi.gene, gene_id, .keep_all = TRUE)
hits <- hits %>% group_by(Epi.gene) %>% distinct(sstart, ssend, .keep_all = TRUE)

levels(as.factor(hits$queryprot))

countsx <- as.data.frame(assay(gvst_dev))
gene.counts <- countsx[(row.names(countsx) %in% hits$gene_id), ]
#str(gene.counts)
gene.counts$gene_id <- row.names(gene.counts) 
gene.counts <- merge(gene.counts, hits[,c(1,3)], by="gene_id")

gene.counts.long <- pivot_longer(gene.counts, 2:25, values_to = "exp")
  #gather(gene.counts, group, exp, X1101:X375, factor_key=TRUE) 
colnames(gene.counts.long)[3] <- "sample_id"
gene.counts.long <- merge(gene.counts.long, sample.info[,c(1:2)], by="sample_id")
#str(gene.counts.long)

mean.counts <- gene.counts.long %>%
  group_by(gene_id, Epi.gene, time_point) %>%
  summarise(mean = mean(exp),
            sd=sd(exp))

mean.counts$time_point <- factor(mean.counts$time_point, 
                                 levels = c("Unfertilized_egg","Fertilized_egg","Cleavage", 
                                  "Prawn_chip","Early_gastrula","Mid_gastrula", "Late_gastrula",
                                  "Planula", "Adult"))
levels(as.factor(mean.counts$Epi.gene))
```

```{r}
DNMT3.counts <- subset(mean.counts, Epi.gene =="DNMT3A")
nDNMT3 <- length(levels(as.factor(DNMT3.counts$gene_id)))
DNMT1.counts <- subset(mean.counts, Epi.gene =="DNMT1")
nDNMT1 <- length(levels(as.factor(DNMT1.counts$gene_id)))
TET1.counts <- subset(mean.counts, Epi.gene =="TET1")
nTET <- length(levels(as.factor(TET1.counts$gene_id)))
MBD2.counts <- subset(mean.counts, Epi.gene =="MBD2")
nMBD2 <- length(levels(as.factor(MBD2.counts$gene_id)))
MBD3.counts <- subset(mean.counts, Epi.gene =="MBD3")
nMBD3 <- length(levels(as.factor(MBD3.counts$gene_id)))
UHRF1.counts <- subset(mean.counts, Epi.gene =="UHRF1")
nUHRF1 <- length(levels(as.factor(UHRF1.counts$gene_id)))
BRG1.counts <- subset(mean.counts, Epi.gene =="BRG1")
nBRG1 <- length(levels(as.factor(BRG1.counts$gene_id)))

SMG.counts <- subset(mean.counts, Epi.gene =="SMG") #SMAUG
nSMG <- length(levels(as.factor(SMG.counts$gene_id)))
KAISO.counts <- subset(mean.counts, Epi.gene =="KAISO") #Kaiso
nKAISO <- length(levels(as.factor(KAISO.counts$gene_id)))
TBXT.counts <- subset(mean.counts, Epi.gene =="TBXT") #brachyury
nTBXT <- length(levels(as.factor(TBXT.counts$gene_id)))
SOX2.counts <- subset(mean.counts, Epi.gene =="SOX2") #SOX2
nSOX2 <- length(levels(as.factor(SOX2.counts$gene_id)))
CyclinB.counts <- subset(mean.counts, Epi.gene =="CyclinB") #CyclinB
nCyclinB <- length(levels(as.factor(CyclinB.counts$gene_id)))
WNT8.counts <- subset(mean.counts, Epi.gene =="WNT8") #WNT8
nWNT8 <- length(levels(as.factor(WNT8.counts$gene_id)))

nEpihomologues <- as.data.frame(cbind(DNMT3A=nDNMT3, DNMT1=nDNMT1, TET1=nTET, MBD2=nMBD2, MBD3=nMBD3, UHRF1=nUHRF1, BRG1=nBRG1))
print(nEpihomologues)
nMZThomologues <- as.data.frame(cbind(SMG=nSMG, KAISO=nKAISO, TBXT=nTBXT, SOX2=nSOX2, CyclinB=nCyclinB, WNT8=nWNT8))
print(nMZThomologues)
```

Select only the top hit for each enzyme for plotting.
```{r}
top_DNMT1=c("g53952")
top_DNMT3=c("g25804")
#skipping TET1 and MBD proteins because they only have 1 hit
top_UHRF1=c("adi2mcaRNA19502_R1")
top_BRG1=c("g68733")
top_SMG=c("g4639")
top_CyclinB=c("g71356")
top_KAISO=c("g60350")
top_TBXT=c("g68947")
top_SOX2=c("g53225")
top_WNT8=c("g33149")
```

## Plot methylation-related genes

Plot DNMT3a expression as a function of time and treatment
```{r}
Mcap.DNMT3 <- DNMT3.counts %>% subset(gene_id%in%top_DNMT3) %>% 
  ggplot(aes(x=time_point, y=mean, group=gene_id)) + 
  geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd), width=.1) +
  geom_line(aes(), lty="dashed", color = "darkgrey") +
  geom_point(aes(), size = 1.5) +
  xlab("Timepoint") +
  ylab(expression(paste("Mean expression (vst-transformed counts)"))) +
  theme_bw() + #Set the background color
  ylim(8.728599, 14) +
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        plot.background=element_blank(),  #Set the plot background
        axis.title.x=element_blank(), #set x-axis title size
        axis.text.x=element_blank(), #set x-axis label size
        axis.text.y=element_text(size = 6), #set x-axis label size
        axis.title.y=element_text(size = 6), #set x-axis label size
        axis.ticks.x=element_blank(), #No x-label ticks
        legend.text=element_text(size = 6), #Legend text size
        legend.title=element_text(size = 6), #Legend title size
        legend.position=c(0.75, 0.9)) +
  ggtitle("i) DNMT3A") +
  #facet_grid(rows=vars(gene_id), scales="free_y") +
  theme(strip.text.y = element_text(size = 5)) +
  theme(plot.title = element_text(face = 'bold.italic', 
                                  size = 9, 
                                  hjust = 0)) +
  annotate("text", x = 3, y = 14, label = "de novo methylation", size = 2.5)
Mcap.DNMT3
```

Plot DNMT1 expression as a function of time and treatment
```{r}
Mcap.DNMT1 <- DNMT1.counts %>% subset(gene_id%in%top_DNMT1) %>% 
  ggplot(aes(x=time_point, y=mean, group=gene_id)) + 
  geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd), width=.1) +
  geom_line(aes(), lty="dashed", color = "darkgrey") +
  geom_point(aes(), size = 1.5) +
  xlab("Timepoint") +
  ylab(expression(paste("Mean normalized expression"))) +
  theme_bw() + #Set the background color
  ylim(8.728599, 14) +
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        plot.background=element_blank(),  #Set the plot background
        axis.title.x=element_blank(), #set x-axis title size
        axis.text.x=element_blank(), #set x-axis label size
        axis.text.y=element_text(size = 6), #set x-axis label size
        axis.title.y=element_blank(), #set x-axis label size
        axis.ticks.x=element_blank(), #No x-label ticks
        legend.text=element_text(size = 6), #Legend text size
        legend.title=element_text(size = 6), #Legend title size
        legend.position=c(0.75, 0.9)) +
  ggtitle("ii) DNMT1") +
  #facet_grid(rows=vars(gene_id), scales="free_y") +
  theme(strip.text.y = element_text(size = 5)) +
  theme(plot.title = element_text(face = 'bold.italic', 
                                  size = 9, 
                                  hjust = 0)) +
  annotate("text", x = 3.5, y = 14, label = "methylation maintenance", size = 2.5)
Mcap.DNMT1
```

Plot TET1 expression as a function of time and treatment
```{r}
Mcap.TET1 <- ggplot(TET1.counts, aes(x=time_point, y=mean, group=gene_id)) + 
  geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd), width=.1) +
  geom_line(aes(), lty="dashed", color = "darkgrey") +
  geom_point(aes(), size = 1.5) +
  xlab("Timepoint") +
  ylab(expression(paste("Mean expression (vst-transformed counts)"))) +
  theme_bw() + #Set the background color
  ylim(8.728599, 14) +
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        plot.background=element_blank(),  #Set the plot background
        axis.title.x=element_blank(), #set x-axis title size
        axis.text.x=element_blank(), #set x-axis label size
        axis.text.y=element_text(size = 6), #set x-axis label size
        axis.title.y=element_blank(), #set x-axis label size
        axis.ticks.x=element_blank(), #No x-label ticks
        legend.text=element_text(size = 6), #Legend text size
        legend.title=element_text(size = 6), #Legend title size
        legend.position=c(0.75, 0.9)) +
  ggtitle("iii) TET1") +
  #facet_grid(rows=vars(gene_id), scales="free_y") +
  theme(strip.text.y = element_text(size = 5)) +
  theme(plot.title = element_text(face = 'bold.italic', 
                                  size = 9, 
                                  hjust = 0)) +
  annotate("text", x = 3, y = 14, label = "methylation removal", size = 2.5)
Mcap.TET1
```

Plot MBD2 and MBD3 expression as a function of time and treatment
```{r}
Mcap.MBD2 <- MBD2.counts %>%
  ggplot(aes(x=time_point, y=mean, group=gene_id)) + 
  geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd), width=.1) +
  geom_line(lty="dashed", color = "darkgrey") +
  geom_point(aes(), size = 1.5) +
  xlab("Timepoint") +
  ylab(expression(paste("Mean expression (vst-transformed counts)"))) +
  theme_bw() + #Set the background color
  ylim(8.728599, 14) +
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        plot.background=element_blank(),  #Set the plot background
        axis.title.x=element_text(size = 6), #set x-axis title size
        axis.text.x=element_text(angle = 90, vjust = 0.5, hjust=1, size = 6), #set x-axis label size
        axis.text.y=element_text(size = 6), #set x-axis label size
        axis.title.y=element_text(size = 6), #set x-axis label size
        axis.ticks.x=element_blank(), #No x-label ticks
        legend.text=element_text(size = 6), #Legend text size
        legend.title=element_text(size = 6), #Legend title size
        legend.position=c(0.75, 0.9)) +
  ggtitle("iv) MBD2/3") +
  #facet_grid(rows=vars(gene_id), scales="free_y") +
  theme(strip.text.y = element_text(size = 5)) +
  theme(plot.title = element_text(face = 'bold.italic', 
                                  size = 9, 
                                  hjust = 0)) +
  annotate("text", x = 3.5, y = 14, label = "transcriptional repression", size = 2.5)
Mcap.MBD2
```

Plot UHRF1 expression as a function of time and treatment
```{r}
Mcap.UHRF1 <- UHRF1.counts %>% subset(gene_id%in%top_UHRF1) %>% ggplot(aes(x=time_point, y=mean, group=gene_id)) + 
  geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd), width=.1) +
  geom_line(lty="dashed", color = "darkgrey") +
  geom_point(aes(), size = 1.5) +
  xlab("Timepoint") +
  ylab(expression(paste("Mean expression (vst-transformed counts)"))) +
  theme_bw() + #Set the background color
  ylim(8.728599, 14.5) +
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        plot.background=element_blank(),  #Set the plot background
        axis.title.x=element_text(size = 6), #set x-axis title size
        axis.text.x=element_text(angle = 90, vjust = 0.5, hjust=1, size = 6), #set x-axis label size
        axis.text.y=element_text(size = 6), #set x-axis label size
        axis.title.y=element_blank(), #set x-axis label size
        axis.ticks.x=element_blank(), #No x-label ticks
        legend.text=element_text(size = 6), #Legend text size
        legend.title=element_text(size = 6), #Legend title size
        legend.position=c(0.75, 0.85)) +
  ggtitle("v) UHRF1") +
  #facet_grid(rows=vars(gene_id), scales="free_y") +
  theme(strip.text.y = element_text(size = 5)) +
  theme(plot.title = element_text(face = 'bold.italic', 
                                  size = 9, 
                                  hjust = 0)) +
  annotate("text", x = 3.5, y = 14.5, label = "methylation maintenance", size = 2.5)
Mcap.UHRF1
```

Plot BRG1 expression as a function of time and treatment
```{r}
Mcap.BRG1 <- BRG1.counts %>% subset(gene_id%in%top_BRG1) %>% 
  ggplot(aes(x=time_point, y=mean, group=gene_id)) +
  geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd), width=.1) +
  geom_line(lty="dashed", color = "darkgrey") +
  geom_point(aes(), size = 1.5) +
  xlab("Timepoint") +
  ylab(expression(paste("Mean expression (vst-transformed counts)"))) +
  theme_bw() + #Set the background color
  ylim(8.728599, 14) +
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        plot.background=element_blank(),  #Set the plot background
        axis.title.x=element_text(size = 6), #set x-axis title size
        axis.text.x=element_text(angle = 90, vjust = 0.5, hjust=1, size = 6), #set x-axis label size
        axis.text.y=element_text(size = 6), #set x-axis label size
        axis.title.y=element_blank(), #set x-axis label size
        axis.ticks.x=element_blank(), #No x-label ticks
        legend.text=element_text(size = 6), #Legend text size
        legend.title=element_text(size = 6), #Legend title size
        legend.position = c(0.75, 0.9)) +
  #guides(shape=guide_legend(ncol=2)) +
  ggtitle("vi) BRG1") +
  #facet_grid(rows=vars(gene_id), scales="free_y") +
  theme(strip.text.y = element_text(size = 5)) +
  theme(plot.title = element_text(face = 'bold.italic',
                                  size = 9,
                                  hjust = 0)) +
  annotate("text", x = 3.25, y = 14, label = "chromatin accessibility", size = 2.5)
Mcap.BRG1
```

Compile plots into single figure and save
```{r}
Fig.7 <- cowplot::plot_grid(Mcap.DNMT3, Mcap.DNMT1, Mcap.TET1, Mcap.MBD2, Mcap.UHRF1, Mcap.BRG1, align = "vh", ncol = 3)
Fig.7
ggsave("2b-Epi-MZT-Enzyme-Expression/Output/Fig7-Mcap-Dev-Epi-Expression.pdf", Fig.7, width=7, height=5)
```

## Plot MZT biomarkers

Plot CyclinB expression as a function of time and treatment
```{r}
Mcap.CyclinB <- CyclinB.counts %>% subset(gene_id%in%top_CyclinB) %>% 
  ggplot(aes(x=time_point, y=mean, group=gene_id)) + 
  geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd), width=.1) +
  geom_line(lty="dashed", color = "darkgrey") +
  geom_point(aes(), size = 1.5) +
  xlab("Timepoint") +
  ylab(expression(paste("Mean expression (vst-transformed counts)"))) +
  theme_bw() + #Set the background color
  #ylim(8,19) +
    theme(axis.line = element_line(color = 'black'), #Set the axes color
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        plot.background=element_blank(),  #Set the plot background
        axis.title.x=element_blank(), #set x-axis title size
        axis.text.x=element_blank(), #set x-axis label size
        axis.text.y=element_text(size = 6), #set x-axis label size
        axis.title.y=element_text(size = 6), #set x-axis label size
        axis.ticks.x=element_blank(), #No x-label ticks
        legend.title=element_text(size = 6), #Legend text size
        legend.text=element_text(size = 6), #Legend text size
        legend.position = c(0.75, 0.9)) +
  ggtitle("i) CyclinB") +
  #facet_grid(rows=vars(gene_id), scales="free_y") +
  theme(strip.text.y = element_text(size = 5)) +
  theme(plot.title = element_text(face = 'bold.italic', 
                                  size = 9, 
                                  hjust = 0)) +
  annotate("text", x = 3.5, y = 19.5, label = "cell-cycle regulation (M/Z)", size = 2.5)
Mcap.CyclinB
```

Plot SMAUG expression as a function of time and treatment
```{r}
Mcap.SMG <- SMG.counts %>% subset(gene_id%in%top_SMG) %>% ggplot(aes(x=time_point, y=mean, group=gene_id)) + 
  geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd), width=.1) +
  geom_line(aes(), lty="dashed", color = "darkgrey") +
  geom_point(aes(), size = 1.5) +
  xlab("Timepoint") +
  ylab(expression(paste("Mean expression (vst-transformed counts)"))) +
  theme_bw() + #Set the background color
  ylim(8.728599, 15.5) +
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        plot.background=element_blank(),  #Set the plot background
        axis.title.x=element_blank(), #set x-axis title size
        axis.text.x=element_blank(), #set x-axis label size
        axis.text.y=element_text(size = 6), #set x-axis label size
        axis.title.y=element_blank(), #set x-axis label size
        axis.ticks.x=element_blank(), #No x-label ticks
        legend.text=element_text(size = 6), #Legend text size
        legend.title=element_text(size = 6), #Legend text size
        legend.position = c(0.85, 0.9)) +
  ggtitle("ii) SMAUG") +
  #facet_grid(rows=vars(gene_id), scales="free_y") +
  theme(strip.text.y = element_text(size = 5)) +
  theme(plot.title = element_text(face = 'bold.italic', 
                                  size = 9, 
                                  hjust = 0)) +
  annotate("text", x = 4.5, y = 15.5, label = "maternal mRNA degradation (M/Z)", size = 2.5)
Mcap.SMG
```

Plot KAISO expression as a function of time and treatment
```{r}
Mcap.KAISO <- KAISO.counts %>% subset(gene_id%in%top_KAISO) %>% 
  ggplot(aes(x=time_point, y=mean, group=gene_id)) + 
  geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd), width=.1) +
  geom_line(aes(), lty="dashed", color = "darkgrey") +
  geom_point(aes(), size = 1.5) +
  xlab("Timepoint") +
  ylab(expression(paste("Mean expression (vst-transformed counts)"))) +
  theme_bw() + #Set the background color
  ylim(8.728599, 15.5) +
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        plot.background=element_blank(),  #Set the plot background
        axis.title.x=element_blank(), #set x-axis title size
        axis.text.x=element_blank(), #set x-axis label size
        axis.text.y=element_text(size = 6), #set x-axis label size
        axis.title.y=element_blank(), #set x-axis label size
        axis.ticks.x=element_blank(), #No x-label ticks
        legend.text=element_text(size = 6), #Legend text size
        legend.title=element_text(size = 6), #Legend text size
        legend.position = c(0.75, 0.9)) +
  ggtitle("iii) KAISO") +
  #facet_grid(rows=vars(gene_id), scales="free_y") +
  theme(strip.text.y = element_text(size = 5)) +
  theme(plot.title = element_text(face = 'bold.italic', 
                                  size = 9, 
                                  hjust = 0)) +
  annotate("text", x = 3.5, y = 15.5, label = "repression of ZGA (M/Z)", size = 2.5)
Mcap.KAISO
```

Plot SOX2 expression as a function of time and treatment
```{r}
Mcap.SOX2 <- SOX2.counts %>% subset(gene_id%in%top_SOX2) %>% 
  ggplot(aes(x=time_point, y=mean, group=gene_id)) +
  geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd), width=.1) +
  geom_line(lty="dashed", color = "darkgrey") +
  geom_point(aes(), size = 1.5) +
  xlab("Timepoint") +
  ylab(expression(paste("Mean expression (vst-transformed counts)"))) +
  theme_bw() + #Set the background color
  ylim(8.728599, 15.5) +
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        plot.background=element_blank(),  #Set the plot background
        axis.title.x=element_text(size = 6), #set x-axis title size
        axis.text.x=element_text(angle = 90, vjust = 0.5, hjust=1, size = 6), #set x-axis label size
        axis.text.y=element_text(size = 6), #set x-axis label size
        axis.title.y=element_text(size = 6), #set y-axis title size
        axis.ticks.x=element_blank(), #No x-label ticks
        legend.text=element_text(size = 6), #Legend text size
        legend.title=element_text(size = 6), #Legend text size
        legend.position = c(0.75, 0.9)) +
  ggtitle("iv) SOX2") +
  #facet_grid(rows=vars(gene_id), scales="free_y") +
  theme(strip.text.y = element_text(size = 5)) +
  theme(plot.title = element_text(face = 'bold.italic', 
                                  size = 9, 
                                  hjust = 0)) +
  annotate("text", x = 2.25, y = 15.5, label = "ZGA (M/Z)", size = 2.5)
Mcap.SOX2
```

Plot WNT8 expression as a function of time and treatment
```{r}
Mcap.WNT8 <- WNT8.counts %>% subset(gene_id%in%top_WNT8) %>% 
  ggplot(aes(x=time_point, y=mean, group=gene_id)) + 
  geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd), width=.1) +
  geom_line(lty="dashed", color = "darkgrey") +
  geom_point(aes(), size = 1.5) +
  xlab("Timepoint") +
  ylab(expression(paste("Mean expression (vst-transformed counts)"))) +
  theme_bw() + #Set the background color
  ylim(8.728599, 15.5) +
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        plot.background=element_blank(),  #Set the plot background
        axis.title.x=element_text(size = 6), #set x-axis title size
        axis.text.x=element_text(angle = 90, vjust = 0.5, hjust=1, size = 6), #set x-axis label size
        axis.text.y=element_text(size = 6), #set x-axis label size
        axis.title.y=element_blank(), #set x-axis label size
        axis.ticks.x=element_blank(), #No x-label ticks
        legend.text=element_text(size = 6), #Legend text size
        legend.title=element_text(size = 6), #Legend text size
        legend.position = c(0.75, 0.9),
        legend.spacing.y = unit(0.01, 'cm')) +
  #guides(shape=guide_legend(ncol=2)) +
  ggtitle("v) WNT8") +
  #facet_grid(rows=vars(gene_id), scales="free_y") +
  theme(strip.text.y = element_text(size = 5)) +
  theme(plot.title = element_text(face = 'bold.italic', 
                                  size = 9, 
                                  hjust = 0)) +
  annotate("text", x = 3.5, y = 15.5, label = "bauplan formation (M/Z)", size = 2.5)
Mcap.WNT8
```

Plot TBXT expression as a function of time and treatment
```{r}
Mcap.TBXT <- TBXT.counts %>% subset(gene_id%in%top_TBXT) %>% 
  ggplot(aes(x=time_point, y=mean, group=gene_id)) + 
  geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd), width=.1) +
  geom_line(lty="dashed", color = "darkgrey") +
  geom_point(aes(), size = 1.5) +
  xlab("Timepoint") +
  ylab(expression(paste("Mean expression (vst-transformed counts)"))) +
  theme_bw() + #Set the background color
  ylim(8.728599, 15.5) +
    theme(axis.line = element_line(color = 'black'), #Set the axes color
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        plot.background=element_blank(),  #Set the plot background
        axis.title.x=element_text(size = 6), #set x-axis title size
        axis.text.x=element_text(angle = 90, vjust = 0.5, hjust=1, size = 6), #set x-axis label size
        axis.text.y=element_text(size = 6), #set x-axis label size
        axis.title.y=element_text(size = 6), #set x-axis label size
        axis.ticks.x=element_blank(), #No x-label ticks
        legend.text=element_text(size = 6), #Legend text size
        legend.title=element_blank(), #Legend text size
        legend.position = c(0.85, 0.85)) +
  ggtitle("vi) TBXT") +
  #facet_grid(rows=vars(gene_id), scales="free_y") +
  theme(strip.text.y = element_text(size = 5)) +
  theme(plot.title = element_text(face = 'bold.italic', 
                                  size = 9, 
                                  hjust = 0)) +
  annotate("text", x = 3.5, y = 15.5, label = "bauplan specification (Z)", size = 2.5)
Mcap.TBXT
```

Compile plots into single figure and save
```{r}
Fig.6 <- cowplot::plot_grid(Mcap.CyclinB, Mcap.SMG, Mcap.KAISO, Mcap.SOX2, Mcap.WNT8, Mcap.TBXT, align = "vh", ncol = 3)
Fig.6
ggsave("2b-Epi-MZT-Enzyme-Expression/Output/Fig6-Mcap-Dev-MZT-Expression.pdf", Fig.6, width=7, height=5)
```

# Run differential expression analysis to see if any biomarkers are differentially expressed
Run differential expression test using a Wald model. 
```{r, message = FALSE}
DEG_dev <- DESeq(gdds_dev) #run differential expression test by group using the Wald model
```

Make a list of biomarkers to look for:
```{r}
Epi.markers <- c(top_DNMT1, top_DNMT3, TET1.counts$gene_id, MBD3.counts$gene_id, top_UHRF1, top_BRG1)

Dev.markers <- c(top_CyclinB, top_SMG, top_KAISO, top_SOX2, top_WNT8, top_TBXT)
```

Look for differential expression of biomarkers
```{r, warning=FALSE}
DESeq.results.fe_vs_ue <-  results(DEG_dev, name="time_point_Fertilized_egg_vs_Unfertilized_egg")
DESeq.results.fe_vs_ue <- as.data.frame(DESeq.results.fe_vs_ue)
DESeq.results.fe_vs_ue$contrast <- c("Fertilized_egg_vs_Unfertilized_egg")
DESeq.results.fe_vs_ue$gene_id <-  rownames(DESeq.results.fe_vs_ue)
fe_vs_ue.DEGs <- filter(DESeq.results.fe_vs_ue, padj<0.05) 
nrow(fe_vs_ue.DEGs) #N DEGs
Epi.DEGs.fe_vs_ue <- filter(DESeq.results.fe_vs_ue, gene_id%in%Epi.markers, padj<0.05, log2FoldChange>1 | log2FoldChange<(-1))
Dev.DEGs.fe_vs_ue <- filter(DESeq.results.fe_vs_ue, gene_id%in%Dev.markers, padj<0.05, log2FoldChange>1 | log2FoldChange<(-1))

DESeq.results.clvg_vs_fe <-  results(DEG_dev, contrast= c("time_point","Cleavage","Fertilized_egg"))
DESeq.results.clvg_vs_fe <- as.data.frame(DESeq.results.clvg_vs_fe)
DESeq.results.clvg_vs_fe$contrast <- c("Cleavage_vs_Fertilized_egg")
DESeq.results.clvg_vs_fe$gene_id <-  rownames(DESeq.results.clvg_vs_fe)
clvg_vs_fe.DEGs <- filter(DESeq.results.clvg_vs_fe, padj<0.05) 
nrow(clvg_vs_fe.DEGs) #N DEGs
Epi.DEGs.clvg_vs_fe <- filter(DESeq.results.clvg_vs_fe, gene_id%in%Epi.markers, padj<0.05, log2FoldChange>1 | log2FoldChange<(-1))
Dev.DEGs.clvg_vs_fe <- filter(DESeq.results.clvg_vs_fe, gene_id%in%Dev.markers, padj<0.05, log2FoldChange>1 | log2FoldChange<(-1))

DESeq.results.pc_vs_clvg <-  results(DEG_dev, contrast= c("time_point","Prawn_chip","Cleavage"))
DESeq.results.pc_vs_clvg <- as.data.frame(DESeq.results.pc_vs_clvg)
DESeq.results.pc_vs_clvg$contrast <- c("Prawn_chip_vs_Cleavage")
DESeq.results.pc_vs_clvg$gene_id <-  rownames(DESeq.results.pc_vs_clvg)
pc_vs_clvg.DEGs <- filter(DESeq.results.pc_vs_clvg, padj<0.05) 
nrow(pc_vs_clvg.DEGs) #N DEGs
Epi.DEGs.pc_vs_clvg <- filter(DESeq.results.pc_vs_clvg, gene_id%in%Epi.markers, padj<0.05, log2FoldChange>1 | log2FoldChange<(-1))
Dev.DEGs.pc_vs_clvg <- filter(DESeq.results.pc_vs_clvg, gene_id%in%Dev.markers, padj<0.05, log2FoldChange>1 | log2FoldChange<(-1))

DESeq.results.eg_vs_pc <-  results(DEG_dev, contrast= c("time_point", "Early_gastrula","Prawn_chip"))
DESeq.results.eg_vs_pc <- as.data.frame(DESeq.results.eg_vs_pc)
DESeq.results.eg_vs_pc$contrast <- c("Early_gastrula_vs_Prawn_chip")
DESeq.results.eg_vs_pc$gene_id <-  rownames(DESeq.results.eg_vs_pc)
eg_vs_pc.DEGs <- filter(DESeq.results.eg_vs_pc, padj<0.05) 
nrow(eg_vs_pc.DEGs) #N DEGs
Epi.DEGs.eg_vs_pc <- filter(DESeq.results.eg_vs_pc, gene_id%in%Epi.markers, padj<0.05, log2FoldChange>1 | log2FoldChange<(-1))
Dev.DEGs.eg_vs_pc <- filter(DESeq.results.eg_vs_pc, gene_id%in%Dev.markers, padj<0.05, log2FoldChange>1 | log2FoldChange<(-1))

DESeq.results.mg_vs_eg <-  results(DEG_dev, contrast= c("time_point", "Mid_gastrula", "Early_gastrula"))
DESeq.results.mg_vs_eg <- as.data.frame(DESeq.results.mg_vs_eg)
DESeq.results.mg_vs_eg$contrast <- c("Mid_gastrula_vs_Early_gastrula")
DESeq.results.mg_vs_eg$gene_id <-  rownames(DESeq.results.mg_vs_eg)
mg_vs_eg.DEGs <- filter(DESeq.results.mg_vs_eg, padj<0.05) 
nrow(mg_vs_eg.DEGs) #N DEGs
Epi.DEGs.mg_vs_eg <- filter(DESeq.results.mg_vs_eg, gene_id%in%Epi.markers, padj<0.05, log2FoldChange>1 | log2FoldChange<(-1))
Dev.DEGs.mg_vs_eg <- filter(DESeq.results.mg_vs_eg, gene_id%in%Dev.markers, padj<0.05, log2FoldChange>1 | log2FoldChange<(-1))

DESeq.results.lg_vs_mg <-  results(DEG_dev, contrast= c("time_point","Late_gastrula","Mid_gastrula"))
DESeq.results.lg_vs_mg <- as.data.frame(DESeq.results.lg_vs_mg)
DESeq.results.lg_vs_mg$contrast <- c("Late_gastrula_vs_Mid_gastrula")
DESeq.results.lg_vs_mg$gene_id <-  rownames(DESeq.results.lg_vs_mg)
lg_vs_mg.DEGs <- filter(DESeq.results.lg_vs_mg, padj<0.05) 
nrow(lg_vs_mg.DEGs) #N DEGs
Epi.DEGs.lg_vs_mg <- filter(DESeq.results.lg_vs_mg, gene_id%in%Epi.markers, padj<0.05, log2FoldChange>1 | log2FoldChange<(-1))
Dev.DEGs.lg_vs_mg <- filter(DESeq.results.lg_vs_mg, gene_id%in%Dev.markers, padj<0.05, log2FoldChange>1 | log2FoldChange<(-1))

DESeq.results.pln_vs_lg <-  results(DEG_dev, contrast= c("time_point", "Planula", "Late_gastrula"))
DESeq.results.pln_vs_lg <- as.data.frame(DESeq.results.pln_vs_lg)
DESeq.results.pln_vs_lg$contrast <- c("Planula_vs_Late_gastrula")
DESeq.results.pln_vs_lg$gene_id <-  rownames(DESeq.results.pln_vs_lg)
pln_vs_lg.DEGs <- filter(DESeq.results.pln_vs_lg, padj<0.05) 
nrow(pln_vs_lg.DEGs) #N DEGs
Epi.DEGs.pln_vs_lg <- filter(DESeq.results.pln_vs_lg, gene_id%in%Epi.markers, padj<0.05, log2FoldChange>1 | log2FoldChange<(-1))
Dev.DEGs.pln_vs_lg <- filter(DESeq.results.pln_vs_lg, gene_id%in%Dev.markers, padj<0.05, log2FoldChange>1 | log2FoldChange<(-1))

DESeq.results.adult_vs_pln <-  results(DEG_dev, contrast= c("time_point", "Adult","Planula"))
DESeq.results.adult_vs_pln <- as.data.frame(DESeq.results.adult_vs_pln)
DESeq.results.adult_vs_pln$contrast <- c("Adult_vs_Planula")
DESeq.results.adult_vs_pln$gene_id <-  rownames(DESeq.results.adult_vs_pln)
adult_vs_pln.DEGs <- filter(DESeq.results.adult_vs_pln, padj<0.05) 
nrow(adult_vs_pln.DEGs) #N DEGs
Epi.DEGs.adult_vs_pln <- filter(DESeq.results.adult_vs_pln, gene_id%in%Epi.markers, padj<0.05, log2FoldChange>1 | log2FoldChange<(-1))
Dev.DEGs.adult_vs_pln <- filter(DESeq.results.adult_vs_pln, gene_id%in%Dev.markers, padj<0.05, log2FoldChange>1 | log2FoldChange<(-1))
```

Bind results. Add enzyme names for biomarker files.
```{r}
#all
all_DEGs <- bind_rows(fe_vs_ue.DEGs, clvg_vs_fe.DEGs, pc_vs_clvg.DEGs, eg_vs_pc.DEGs, mg_vs_eg.DEGs, lg_vs_mg.DEGs, pln_vs_lg.DEGs, adult_vs_pln.DEGs)
rownames(all_DEGs) <- NULL #remove row names
all_DEGs <- all_DEGs[,c(8,1:7)] #move gene_id to first column


#epi
Epi_DEGs <- bind_rows(Epi.DEGs.fe_vs_ue, Epi.DEGs.clvg_vs_fe, Epi.DEGs.pc_vs_clvg, Epi.DEGs.eg_vs_pc, Epi.DEGs.mg_vs_eg, Epi.DEGs.lg_vs_mg, Epi.DEGs.pln_vs_lg, Epi.DEGs.adult_vs_pln)
Epi_DEGs <- merge(Epi_DEGs, hits[,c(1,3)], by="gene_id")
head(Epi_DEGs)

#dev
Dev_DEGs <- bind_rows(Dev.DEGs.fe_vs_ue, Dev.DEGs.clvg_vs_fe, Dev.DEGs.pc_vs_clvg, Dev.DEGs.eg_vs_pc, Dev.DEGs.mg_vs_eg, Dev.DEGs.lg_vs_mg, Dev.DEGs.pln_vs_lg, Dev.DEGs.adult_vs_pln)
Dev_DEGs <- merge(Dev_DEGs, hits[,c(1,3)], by="gene_id")
head(Dev_DEGs)

#save
write_csv(Epi_DEGs, file = "2b-Epi-MZT-Enzyme-Expression/Output/Epi_biomarker_DEGs.csv")
write_csv(Dev_DEGs, file = "2b-Epi-MZT-Enzyme-Expression/Output/Dev_biomarker_DEGs.csv")
#write_csv(all_DEGs, file = "2b-Epi-MZT-Enzyme-Expression/Output/all_DEGs.csv")
```